Please find below the R code history from your *Wallace* session.

You can reproduce your session results by running this R Markdown file in RStudio.

Each code block is called a "chunk", and you can run them either one-by-one or all at once by choosing an option in the "Run" menu at the top-right corner of the "Source" pane in RStudio.

For more detailed information see <http://rmarkdown.rstudio.com>).

### Package installation

Wallace uses the following R packages that must be installed and loaded before starting.

```{r}
library(spocc)
library(spThin)
library(dismo)
library(rgeos)
library(ENMeval)
```

Wallace also includes several functions developed to help integrate different packages and some additional functionality. For this reason, it is necessary to load the file `functions.R`, The function `system.file()` finds this script, and `source()` loads it.

```{r}
source(system.file('shiny/funcs', 'functions.R', package = 'wallace'))
```

Record of analysis for *Lithobates palustris*.
----------------------------------------------

### Obtain Occurrence Data

The search for occurrences was limited to 500 records. Obtain occurrence records of the selected species from the gbif database.

```{r}
# query selected database for occurrence records
results <- spocc::occ(query = "Lithobates palustris", from = "gbif", limit = 500, has_coords = TRUE)
# retrieve data table from spocc object
results.data <- results[["gbif"]]$data[[formatSpName("Lithobates palustris")]]
# remove rows with duplicate coordinates
occs.dups <- duplicated(results.data[c('longitude', 'latitude')])
occs <- results.data[!occs.dups,]
# make sure latitude and longitude are numeric (sometimes they are characters)
occs$latitude <- as.numeric(occs$latitude)
occs$longitude <- as.numeric(occs$longitude)
# give all records a unique ID
occs$occID <- row.names(occs)
# extract occurrence coordinates
occs.xy <- occs[c('longitude', 'latitude')]
```

### Process Occurrence Data

The following code recreates the polygon used to select occurrences to keep in the analysis.

```{r}
selCoords <- data.frame(x = c(-74.3048, -74.7552, -76.7767, -76.8646, -75.3705, -75.0848, -74.3048), y = c(40.3599, 41.6368, 41.9644, 41.8172, 41.0595, 40.2594, 40.3599))
selPoly <- sp::SpatialPolygons(list(sp::Polygons(list(sp::Polygon(selCoords)), ID=1)))
intersect <- sp::over(occs.xy, selPoly)
intersect.rowNums <- as.numeric(which(!(is.na(intersect))))
occs <- occs[intersect.rowNums, ]
```

### Obtain Environmental Data

Using WorldClim (<http://www.worldclim.org/>) bioclimatic dataset at resolution of 2.5 arcmin.

```{r}
# get WorldClim bioclimatic variable rasters
envs <- raster::getData(name = "worldclim", var = "bio", res = 2.5, lat = , lon = )
# subset by those variables selected
envs <- envs[[c('bio01', 'bio02', 'bio03', 'bio04', 'bio05', 'bio06', 'bio07', 'bio08', 'bio09', 'bio10', 'bio11', 'bio12', 'bio13', 'bio14', 'bio15', 'bio16', 'bio17', 'bio18', 'bio19')]]
# extract environmental values at occ grid cells
locs.vals <- raster::extract(envs[[1]], occs[, c('longitude', 'latitude')])
# remove occs without environmental values
occs <- occs[!is.na(locs.vals), ]  
```

### Process Environmental Data

Background selection technique chosen as .

```{r}
# make SpatialPoints object for buffering
occs.sp <- sp::SpatialPoints(occs.xy)
bgExt <- rgeos::gBuffer(occs.sp, width = 0.11)
```

Mask environmental variables by , and take a random sample of background values from the study extent. As the sample is random, your results may be different than those in the session. If there seems to be too much variability in these background samples, try increasing the number from 10,000 to something higher (e.g. 50,000 or 100,000). The better your background sample, the less variability you'll have between runs.

```{r}
# crop the environmental rasters by the background extent shape
envsBgCrop <- raster::crop(envs, bgExt)
# mask the background extent shape from the cropped raster
envsBgMsk <- raster::mask(envsBgCrop, bgExt)
# sample random background points
bg.xy <- dismo::randomPoints(envsBgMsk, 10000)
# convert matrix output to data frame
bg.xy <- as.data.frame(bg.xy)  
```

### Partition Occurrence Data

Occurrence data is now partitioned for cross-validation, a method that iteratively builds a model on all but one group and evaluates that model on the left-out group.

For example, if the data is partitioned into 3 groups A, B, and C, a model is first built with groups A and B and is evaluated on C. This is repeated by building a model with B and C and evaluating on A, and so on until all combinations are done.

Cross-validation operates under the assumption that the groups are independent of each other, which may or may not be a safe assumption for your dataset. Spatial partitioning is one way to ensure more independence between groups.

You selected to partition your occurrence data by the method.

```{r}
group.data <- ENMeval::get.jackknife(occ=occs.xy, bg.coords=bg.xy)
```

```{r}
# pull out the occurrence and background partition group numbers from the list
occs.grp <- group.data[[1]]
bg.grp <- group.data[[2]]
```

### Build and Evaluate Niche Model

You selected the maxent model.

```{r}
# define the vector of regularization multipliers to test
rms <- seq(1, 2, 1)
# iterate model building over all chosen parameter settings
e <- ENMeval::ENMevaluate(occs.xy, envsBgMsk, bg.coords=bg.xy, RMvalues=rms, fc=c('L', 'LQ'), method='user', occs.grp, bg.grp)

# unpack the results data frame, the list of models, and the RasterStack of raw predictions
evalTbl <- e@results
evalMods <- e@models
evalPreds <- e@predictions
```

```{r}
# view ENMeval results
ENMeval::eval.plot(evalTbl, value = "delta.AICc")
```

```{r}
# Select your model from the models list
mod <- evalMods[["L_1"]]
```

```{r}
# generate logistic prediction
pred <- dismo::predict(mod, envsBgMsk, args=c("outputformat=logistic"))
```

```{r}
# get predicted values for occurrence grid cells
occPredVals <- raster::extract(pred, occs.xy)
# define minimum training presence threshold
thr <- thresh(occPredVals, "noThresh")
# threshold model prediction
pred <- pred > thr
```

```{r}
# plot the model prediction
plot(pred)
```

### Project Niche Model

You selected to project your model. First define a polygon with the coordinates you chose, then crop and mask your predictor rasters. Finally, predict suitability values for these new raster cells based on the model you selected.

```{r}
projCoords <- data.frame(x = c(-73.9841, -73.9786, -73.5886, -73.7149, -73.4952, -73.5446, -73.9841), y = c(-73.9841, -73.9786, -73.5886, -73.7149, -73.4952, -73.5446, -73.9841))
projPoly <- sp::SpatialPolygons(list(sp::Polygons(list(sp::Polygon(projCoords)), ID=1)))
```

### Project Niche Model to New Extent

Now use crop and mask the predictor variables by projPoly, and use the predict() function to predict the values for the new extent based on the model selected.

```{r}
predsProj <- raster::crop(envs, projPoly)
predsProj <- raster::mask(predsProj, projPoly)

proj <- dismo::predict(mod, predsProj)
```

```{r}
# get predicted values for occurrence grid cells
occPredVals <- raster::extract(pred, occs.xy)
# define minimum training presence threshold
thr <- thresh(occPredVals, "noThresh")
# threshold model prediction
proj <- proj > thr
```

```{r}
# plot the model prediction
plot(proj)
```
